---
title: "Exploratory Data Analysis"
author: "Andy Wang 305609128"
date: "2024-07-13"
output: pdf_document
---

# Overview

## Files

1. `customer_info_train.csv`, `customer_info_test.csv` - contains customer demographic info
2. `amazon_order_details_train.csv`, `amazon_order_details_test.csv` - contains each order's specific info
3. `train.csv`, `test.csv` - primary data file used for modelling

## Imports

```{r}
library(dplyr)
library(ggplot2)
library(tidymodels)
library(reshape2)
library(lubridate)
```

# File 1: `train.csv`

```{r}
train_path <- "../ucla-stats-101-c-2024-su-regression/train.csv"
train <- read.csv(train_path)
#train <- train %>% select(-order_totals)
head(train)
```

## Variables

index variables
- `q_demos_state`   - US state where orders were placed
- `year`            - year where orders were placed
- `month`           - month where orders were placed

user info
- `count`           - total number of orders placed
- `count_female`    - number of orders placed by those who responded female to survey question
- `count_male`      - number of orders placed by those who responded male to survey question
- `count_less5`     - number of orders placed by those who responded "place less than 5 orders per month on Amazon"
- `count_5to10`     - "5 to 10 orders per month on Amazon"
- `count_over10`    - "over 10 orders per month on Amazon"

household size variables
- `count_hh1`       - household size = 1
- `count_hh2`       - household size = 2
- `count_hh3`       - household size = 3
- `count_hh4`       - household size = 4+

account sharing variables
- `count_howmany1`  - how many people in the household use/share the amazon account = 1
- `count_howmany2`  - how many people in the household use/share the amazon account = 2
- `count_howmany3`  - how many people in the household use/share the amazon account = 3
- `count_howmany4`  - how many people in the household use/share the amazon account = 4

customer age variables
- `count_1824`      - age 18-24
- `count_2534`      
- `count_3544`            
- `count_4554`            
- `count_5564`            
- `count_65up`            

customer income variables
- `count_und25k`    - under 25k
- `count_2549k`     - 25k to 49k
- `count_5074k`           
- `count_7599k`           
- `count_100149k`   - 100k to 149k
- `count_150kup`    - over 150k

customer education variables
- `count_lessHS`    - less than HS
- `count_HS`        - HS diploma
- `count_B`         - Bachelor's degree
- `count_G`         - graduate / professional degree

target variable
- `order_totals`    - total order price
- `log_total`       - log of total order price

## Correlation Matrix

```{r}
correlation_mat <- train %>% 
  select(-order_totals) %>% 
  select_if(is.numeric) %>% 
  cor()

melted_cormat <- melt(correlation_mat)

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

# Use correlation between variables as distance
reorder_cormat <- function(cormat){
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Reorder the correlation matrix
cormat <- reorder_cormat(correlation_mat)
upper_tri <- get_upper_tri(cormat)

# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# Create a heatmap
ggplot(melted_cormat, aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    low = "blue", 
    high = "red", 
    mid = "white", 
    midpoint = 0, 
    limit = c(-1, 1), 
    space = "Lab", 
    name = "Pearson\nCorrelation"
  ) +
  theme_minimal() + # minimal theme
  theme(
    axis.text.x = element_text(angle = 90, size = 9)
  ) +
  coord_fixed()
```

## Distribution Plots

```{r}
# Distribution Plots
# Histogram of log_total
ggplot(train, aes(x = log_total)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of log_total", x = "log_total", y = "Count")
```

## Time Series Plots

```{r}
# Time Series Plots
# Line Plot of log_total over time
timeseries <- train %>%
  group_by(year, month) %>% 
  summarise(log_total = sum(log_total), count = sum(count)) %>% 
  mutate(date = ymd(paste(year, month, "01", sep = "-"))) %>% 
  select(date, log_total, count) %>% 
  ungroup()

ggplot(timeseries, aes(x = date, y = log_total)) +
  geom_line(color = "blue") +
  theme_minimal() +
  labs(title = "log_total Over Time", x = "Date", y = "log_total")

# Line Plot of count over time
ggplot(timeseries, aes(x = date, y = count)) +
  geom_line(color = "green") +
  theme_minimal() +
  labs(title = "Total Orders Over Time", x = "Date", y = "Count")
```

## Geographical Plots

```{r}
# Geographical Plots
# Assuming you have a state map dataset named `us_states` with state abbreviations
us_states <- map_data("state")

states <- train %>%
  mutate(q_demos_state = tolower(q_demos_state)) %>% 
  group_by(q_demos_state) %>%
  summarize(avg_total = mean(order_totals)) %>% 
  select(q_demos_state, avg_total) %>% 
  ungroup()

# Merge the state data
states <- us_states %>%
  left_join(states, by = c("region" = "q_demos_state"))

ggplot(states, aes(map_id = region, fill = avg_total)) +
  geom_map(map = us_states, color = "white") +
  expand_limits(x = us_states$long, y = us_states$lat) +
  theme_minimal() +
  labs(title = "Average order_total by State", fill = "Avg order_totals")
```

## Scatter Plots

```{r}
# Scatter Plots
# Scatter plot of log_total vs count_male
ggplot(train, aes(x = count_male, y = log_total)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(title = "log_total vs Count of Male Orders", x = "Count of Male Orders", y = "log_total")

# Scatter plot of log_total vs count_female
ggplot(train, aes(x = count_female, y = log_total)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(title = "log_total vs Count of Female Orders", x = "Count of Female Orders", y = "log_total")

# Pair Plot
# Using GGally package for a pair plot
library(GGally)
ggpairs(train %>% select(log_total, count_female, count_male, count_1824, count_und25k))
```

## Bar Plots TODO: FIX THIS

```{r}
# Bar Plots
# Bar Plot of average order_totals by age group
age_vars <- c("count_1824", "count_2534", "count_3544", "count_4554", "count_5564", "count_65up")
ages <- train %>%
  select(q_demos_state, year, month, order_totals, all_of(age_vars)) %>%
  pivot_longer(cols = age_vars, names_to = "age_group", values_to = "count")

ages <- ages %>%
  group_by(age_group) %>%
  summarize(avg_total = mean(order_totals))

ggplot(ages, aes(x = age_group, y = avg_total, fill = age_group)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Average order_totals by Age Group", x = "Age Group", y = "Avg order_totals")
```

## Box Plots

```{r}
# Box Plots
# Box Plot of log_total by state
# Calculate the median log_total for each state
state_medians <- train %>%
  group_by(q_demos_state) %>%
  summarize(median_log_total = median(log_total)) %>%
  arrange(median_log_total)

# Reorder the levels of q_demos_state based on the median log_total
state_medians <- train %>%
  mutate(q_demos_state = factor(q_demos_state, levels = state_medians$q_demos_state))

# Box Plot of log_total by state with sorted states
ggplot(state_medians, aes(x = log_total, y = q_demos_state)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Box Plot of log_total by State", y = "State", x = "log_total")

# Box Plot of log_total by month
ggplot(train, aes(x = factor(month), y = log_total)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Box Plot of log_total by Month", x = "Month", y = "log_total")
```





